---
title: |
    | Peer Graded Assignment <br /> R markdown presentation and Plotly
author: "RLC Velhorst"
date: "15-5-2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plotly)
library(dplyr)
set.seed(101)

```


# Sea level rise in the Netherlands | Developing Data Products Assignment Week 3

This document is the report of the Peer Assessment project of week 2 from Coursera’s course Developing Data Products, as part of the Specialization in Data Science by John Hopkins University. It was built in RStudio, using its knitr functions. 

## Project Goal  

Demonstrate skills:  

* R markdown presentation  
* Plotly interactive plot  


## Data: Publicly Available Tide Stations

The specific data in this exercise shows historical sea level in the Netherlands.  

The data used for this assignment originates from the Permanent Service for Mean Sea Level (PSMSL).
PSMSL is the global data bank for long term sea level change information from tide gauges and bottom pressure recorders.

This doument shows a selection of average sea level measurements in the Netherlands, which are stored in de publicly available data: tide gauge data [Holgate et al., 2013; PSMSL, 2020].


```{r, comment=""}

# data directory
datadir <- "./data"
if(!file.exists(datadir)){dir.create(datadir)}

# download zipped data
zipf = file.path(datadir, "rlr_annual.zip")
if(!file.exists(zipf)){
  fileUrl <- "https://www.psmsl.org/data/obtaining/year_end/2019/rlr_annual.zip"
  download.file(fileUrl,destfile = zipf)
  unzip(zipf,exdir="./data")
}

# file containing tide stations locations
fn_data <- "rlr_annual/filelist.txt"
dataf <-file.path(datadir, fn_data)

# station data
stations <- read.table(dataf, sep = ";",
                       col.names = c('id', 'lat', 'lon', 'name', 'coastline_code', 'station_code', 'quality'),
                       strip.white = TRUE
                       )

# Select stations in the Netherlands
NL_coastline_code = 150
stations.NL  <-stations[which(stations$coastline_code==NL_coastline_code),]

head(stations.NL[,c('id', 'lat', 'lon', 'name')], 3)
```
## Dutch Sea Levels

```{r warning=FALSE, message=FALSE, error=FALSE}

# locate sea level data
rlrdata_fp <- function(id){
  # return file path of the sea level data
  file.path('./data/rlr_annual/data', sprintf("%d.rlrdata",id))
}
rlrinfo_fp <- function(id){
  # return file path of the local reference level info
  file.path('./data/rlr_annual/RLR_info', sprintf("%d.txt",id))
}

rlr_correction_df <- function(id){
  # return dataframe containing the RLR correction of a given station (identified by id)
  # correction information is obtained from the rlr info file
  # The dataframe contains three columns: year_start, year_end, and correction [mm]
  
  fp <- rlrinfo_fp(id)
  linestr <- readLines(con <- file(fp))
  close(con)
  records <- strsplit(linestr,split="<br>")
  words <- sapply(records, strsplit,split=" ")
  
  correction_df <- data.frame()
  for(ix in 1:length(words)){
    if(words[[ix]][1]=="Add"){
      if('onwards' %in% words[[ix]]){
        year_start <- as.integer(words[[ix]][6])
        year_end <- as.integer(format(Sys.Date(), "%Y"))
        correction <- as.numeric(strsplit(words[[ix]][2], "m")[1])*1000. # convert to mm
      } else {
        years = strsplit(words[[ix]][6], "-")
        year_start <- as.integer(years[[1]][1])
        year_end <- as.integer(years[[1]][2])
        correction <- as.numeric(strsplit(words[[ix]][2], "m")[1])*1000. # convert to mm
      }
      df_tmp = data.frame("year_start"= year_start,
                          "year_end"= year_end,
                          "correction"= correction)
      correction_df <- rbind(correction_df,df_tmp)
    }
  }
  correction_df
}

getCorrection <- function(year, correction_df) {
  # return the correction of a given year, based on the correction dataframe
  tmp <- correction_df %>%
    filter(year <= year_end, year >= year_start)
  return(tmp$correction)
}

# NAP_convert
rlr2nap <- function(df, id){
  # Add a column "NAP" [mmm] to the sea level dataframe df
  # NAP  is the national reference level in the Netherlands
  # Conversion from RLR to NAP is based on RLR-info of a given station (identified by id)
  correction_df <- rlr_correction_df(id)
  RLR_correction <- sapply(as.vector(df$Year), getCorrection, correction_df)
  return(mutate(df, NAP = RLR - RLR_correction))
}


# Create dataframe of all annual dutch sea levels
# select station id's
NLids = stations.NL$id
# loop through all stations and merge data into one dataframe
RLRdata.NL = data.frame()
for(id in NLids){
  station <- stations.NL[which(id==stations.NL$id),]
  
  df_tmp <- read.table(rlrdata_fp(id), sep =";", col.names= c("Year", "RLR", "Quality", "NUll")) %>% 
      merge(station) %>% # add station information to dataframe
     rlr2nap(id) # Add NAPto dataframe
  
  RLRdata.NL <- rbind(RLRdata.NL,df_tmp)
}

# Compute mean NAP, after 
meanNAP.NL <- aggregate(NAP ~ Year, RLRdata.NL, mean)
# filter out non-trusted part (before NAP)
meanNAP.NL <- filter(meanNAP.NL, Year>=1890)
meanNAP.NL$name= "Mean"

# Combine dataframes
# cols <- c("Year","NAP","name")
# df_combi <- rbind(RLRdata.NL[,cols], meanNAP.NL2[,cols] )
```


```{r}


# fig <- plot_ly(df_combi[which(df_combi$name!="Mean"),], x= ~Year, y= ~NAP,
#         mode = "markers", color = ~name,type="scatter", colors = "Set3") %>%
#         add_trace(df_combi[which(df_combi$name=="Mean"),], x= ~Year, y= ~NAP, mode = 'lines', type='scatter')

# set colors
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "black", "#D55E00", "#CC79A7")

# create plot
# p <- ggplot(data=df_combi[which(df_combi$name!="Mean"),], aes(x = Year, y = NAP, color = name)) +
#   geom_line(data=df_combi[which(df_combi$name=="Mean"),], aes(x = Year, y = NAP)) +

# p <- ggplot(data=RLRdata.NL, aes(x = Year, y = NAP, color = name)) +
#   geom_point(size = 1) +
#   scale_color_brewer(palette="Set3") + 
#   geom_line(data=meanNAP.NL, aes(x = Year, y = NAP)) +
#   scale_colour_manual(values=cbp1)+
#   labs(color = "Station" ) +
#   ylab("Water level [mm] above NAP")+
#   theme_bw()
# 
# # Set plot to plotly
# fig <- ggplotly(p)
# fig %>% layout(legend = list(orientation = "v", x = 1, y = 0))



# create plot
# p <- ggplot(data=df_combi[which(df_combi$name!="Mean"),], aes(x = Year, y = NAP, color = name)) +
#   geom_line(data=df_combi[which(df_combi$name=="Mean"),], aes(x = Year, y = NAP)) +
p <- ggplot(data=RLRdata.NL, aes(x = Year, y = NAP, color = name)) +
  geom_point(size = 1) +
  geom_line(data=meanNAP.NL, aes(x = Year, y = NAP)) +
  scale_colour_manual(values=cbp1)+
  labs(color = "Station" ) +
  ylab("Water level [mm] above NAP")+
  theme_bw() + 
  theme( legend.title = element_blank() )

# Set plot to plotly
fig <- ggplotly(p)
fig %>%
  add_annotations( text="Station", xref="paper", yref="paper",
                  x=1.02, xanchor="left",
                  y=0.8, yanchor="bottom",    # Same y as legend below
                  legendtitle=TRUE, showarrow=FALSE ) %>%
  layout( legend=list(y=0.8, yanchor="top" ) )

```



# Thank you for your attention


